---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# buscadoR

<!-- badges: start -->
[![R-CMD-check](https://github.com/TeamMacLean/buscadoR/workflows/R-CMD-check/badge.svg)](https://github.com/TeamMacLean/buscadoR/actions)
<!-- badges: end -->

The goal of buscadoR is to find plant receptor proteins from provided FASTA sequence. It uses the web API of PFAM and phobius and a local BLAST to find repeat, transmembrane and ecto- domains, signal peptides then applies a heuristic to classify into receptor classes.

## Installation

You can install the development version of buscadoR from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("TeamMacLean/buscadoR")
```

## Performing the web search

To start the searches with a new protein file you need to supply a FASTA file of protein sequences, your valid email address and specify a restart filename for use later. 


```{r example, eval=FALSE}
library(buscadoR)
fasta_sequences <- "true_positive.fa"
saved_progress <- "saved_progress.fa"
searches <- buscar(protein_file = fasta_sequences, restart_file = saved_progress, email="my_name@my.org")
```

The process will run the searches. Phobius searches are really quick. Ectodomain BLASTs are done locally. PFAM can take a while. By default, the process will wait just one second after submission is complete for PFAM before quitting and writing the progress to the `restart_file` you specified. 

In simple cases with just a few (<20) proteins you may wish to wait and to run the searches and get the results in one step. 


In most cases it is expected that the searches are done in two steps. This is the same call but only the `restart_file` is needed, the parameters are carried over from the initial job.
You can check whether the searches have all been retrieved with `completed()` which returns `TRUE` if everything is complete. 


```{r}
searches <- buscar(restart_file = saved_progress)
complete(searches)
```


If you want a progress bar in an interactive session

```{r, eval=FALSE}
searches <- buscar(file_path, email="my_name@my.org", progress=TRUE)
```


And if you need to increase the amount of time the process will wait for the PFAM server before giving up. This example checks every five seconds for 5 minutes

```{r,eval = FALSE}
searches <- buscar(file_path, email="my_name@my.org", progress=TRUE,
                   wait=5, maxchecktime=360 )
```

## Examining the results

A summary table of each type found can be generated

```{r}
mesa(searches)
```

A dataframe of one row per protein (ideal for exporting) can be created

```{r}
as.data.frame(searches)
```

Tidy format dataframe of each type of receptor protein can be extracted 

```{r}
lrr_rp(searches)
#also lrr_rk(), non_lrr_rp(), non_lrr_rk(), lrr_rp_rk_with_ecto()
```

Raw search results from the databases can be extracted

```{r}
pfam_results(searches)
#also phobius_results(), ecto_results()
```

Each set of found proteins can be rendered as a plot

```{r}
dibujar(searches, which = "lrr_rp")
#also use "lrr_rk", "non_lrr_rp", "non_lrr_rk", "lrr_rp_rk_with_ecto"
```

## Exporting results

A dataframe of one row per protein can be created and written out in the usual way

```{r, eval=FALSE}
res <- as.data.frame(searches)
readr::write_csv(res, "my_results.csv")
```

Annotated FASTA sequences can be exported

```{r, eval=FALSE}
write_seqs(searches, "my_seqs.fa")
```

## Other

A dataframe compatible with the `drawProteins` package [here on bioconductor](https://bioconductor.org/packages/release/bioc/vignettes/drawProteins/inst/doc/drawProteins_BiocStyle.html) can be created for further plotting work.

```{r, eval=FALSE}
dp <- as.drawProteins(searches)
```

Note the plot is a `ggplot2` object and can be styled using that package too. 

The definitions used for the classification can be returned

```{r}
definiciones(which="lrr_rp")
#also use "lrr_rk", "non_lrr_rp", "non_lrr_rk", "lrr_rp_rk_with_ecto", "all"
```

